### 2024-03-17 paper/paper.qmd
Implemented suggestions and updated paper

---
title: "Age, Education, Income, and the Canadian Vote. Understanding the 2021 election."
author: Allen Uy
thanks: "Code and data are available at: [https://github.com/varygx/CES2021](https://github.com/varygx/CES2021)"
date: today
date-format: long
abstract: "This study examines the factors shaping support for Canada's major parties, the Liberal and Conservative, during the 2021 federal election. Analyzing personal and topical variables, we discovered that age has minimal impact, higher education is linked to Liberal support, and higher income is linked to Conservative support. These insights offer an understanding of the 2021 election dynamics, enabling more informed predictions for the upcoming 2025 election. This research contributes to a better grasp of Canadian political affiliations, assisting citizens and policymakers in adapting to evolving electoral trends."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library("tidyverse")
library("arrow")
library("here")
library("knitr")
library("kableExtra")

ces21_data <- read_parquet(here::here("data/analysis_data/clean_ces2021.parquet"))
```

# Introduction

The two largest political parties in Canada, the Liberal Party and Conservative party won 160 and 119 out of 338 seats respectively in the 44th federal election held in 2021 (@citeSeats). With the 45th federal election approaching in 2025, this paper conducts an analysis on the political support for these two parties, specifically examining personal and topical variables. The estimand of interest in this research is the average causal effect of these variables on the likelihood of individuals expressing support for either the Liberal or Conservative party.

By exploring personal and topical variables, this paper aims to provide insight into the factors that influenced political support for the Liberal and Conservative parties in Canada. The goal is to provide a clearer understanding of the dynamics that shaped the 2021 election and offer insights into broader patterns of political affiliations in the Canadian context, which could be used for inference in the upcoming election.

Our findings show that age has little to no effect on party support and those with higher education are likely to vote Liberal while those with higher income are more likely to vote Conservative.

The subsequent sections follow a structured format. @sec-data outlines the source and variables of interest for our analysis. @sec-model details the construction and methodology of the statistical models used. @sec-results presents the key findings of our analysis, while @sec-discussion critically reviews the content, addresses the implications of the results, acknowledges model limitations, and suggests potential research directions.

# Data {#sec-data}

The data used in this paper was gathered from the 2021 Canadian Election Study (CES) hosted on the Harvard Dataverse [@citeCES] and analyzed using R [@citeR] with help from `tidyverse` [@citeTidyverse], `haven` [@citeHaven], `arrow` [@citeArrow], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `here` [@citeHere], `knitr` [@citeKnitr], and `kableExtra` [@citeKableExtra].

## CES2021

The dataset was gathered by CES through an online survey via the Leger Opinion platform where data was finalized on November 1, 2023. A total of 20968 responses were recorded with the stratification goals contained in [Appendix -@sec-stratification-goals].

The survey was launched during what they defined as the Campaign Period Survey (CPS) during August 17 to September 19, 2021 and a follow-up Post-Election Survey (PES) had 15069 responses during September 23 to October 4, 2021.

The provided data was cleaned by CES to include mostly high-quality responses, the details of which can be found in [Appendix -@sec-response-quality].

The dataset includes 1059 variables, many of which could have been included in the analysis but was narrowed down to 30 variables that could correlate with party support.

These variables are gathered from the CPS portion of the survey with no open-ended answers and assigned numerical values with labels.

### Personal Variables

Age was calculated in years based on the respondent's year of birth. Education, Employment, Religion, Immigration Status, and Province were given a number with a corresponding label based on which radio button the respondent selected. Income category was given a number corresponding to a range of household incomes. Respondents were asked to input their income or if they felt uncomfortable to only specify the range. Exact income numbers were converted to their appropriate range. A preview of personal variables can be seen in @tbl-personal_var.

```{r, echo=FALSE}
#| label: tbl-personal_var
#| tbl-cap: Sample of Personal Variable Data
head(ces21_data, 5) %>% 
  select(age, education, income_cat) %>% 
  kable(
    booktabs = TRUE,
    linesep = ""
  ) %>% kable_styling(latex_options = "scale_down")
```

```{r, echo=FALSE}
#| label: fig-age_summary
#| fig-cap: Distribution of respondent age
ces21_data %>% ggplot(aes(x=age)) +
  geom_bar() +
  theme_minimal()
```

```{r, echo=FALSE}
#| label: fig-education_summary
#| fig-cap: Distribution of respondent education level
ces21_data %>% mutate(education = recode(education,
                                        "Some elementary school" = "Some elementary",
                                        "Completed elementary school" = "Completed elementary",
                                        "Some secondary/ high school" = "Some secondary",
                                        "Completed secondary/ high schoo" = "Completed secondary", 
                                        "Some technical, community college, CEGEP, College Classique" = "Some college",
                                        "Completed technical, community college, CEGEP, College Classique" = "Completed college")) %>% ggplot(aes(x=education)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
#| label: fig-income_summary
#| fig-cap: Distribution of respondent income level

ces21_data %>% ggplot(aes(x=income_cat)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_discrete(limit = c("No income", "$1 to $30,000", "$30,001 to $60,000",
                       "$60,001 to $90,000", "$90,001 to $110,000", "$110,001 to $150,000",
                       "$150,001 to $200,000", "More than $200,000", "Don't know/ Prefer not to answer"))
```

The age distribution seen in @fig-age_summary matches voter turnout by age group in previous elections [@citeAgeTurnout]. @fig-education_summary shows the majority of respondents have completed college or university. The majority of respondents are close to the median total income for individuals and families in 2021 [@citeMedianIncome].

### Topic Variables

The following survey topics were extracted from the dataset: Economy, Healthcare, Immigration, COVID, Environment, Quebec Sovereignty, Government Spending, Education, Housing, Carbon Tax, and Childcare. Variables ending in `_party` indicate the party which the respondent believes would handle the best. The survey was designed to give only half the respondents this question. Other variables are from questions with responses on a scale such as: "(strongly) disagree, neutral, (strongly) agree" or "spend less, spend the same, spend more" where the respondent selected one of those answers or did not answer. A preview of topic variables can be seen in @tbl-topic_var.

```{r, echo=FALSE}
#| label: tbl-topic_var
#| tbl-cap: Sample of Topic Variable Data
head(ces21_data, 5) %>% 
  select(econ_party, educ_party, spend_educ) %>% 
  kable(
    booktabs = TRUE,
    linesep = ""
  ) %>% kable_styling(latex_options = "scale_down")
```

```{r, echo=FALSE}
#| label: fig-econ_party_summary
#| fig-cap: Responses to "Which party would do the best job at handling economy?"

ces21_data %>% ggplot(aes(x=econ_party)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
#| label: fig-educ_party_summary
#| fig-cap: Responses to "Which party would do the best job at handling education?"

ces21_data %>% ggplot(aes(x=educ_party)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
#| label: fig-educ_spend_summary
#| fig-cap: Responses to "How much should the federal government spend on education?"

ces21_data %>% ggplot(aes(x=spend_educ)) +
  geom_bar() +
  theme_minimal()
```

@fig-econ_party_summary shows that more respondents believe the Conservative party would handle economy the best out of all parties while @fig-educ_party_summary shows that more respondents believe the Liberal party would handle education the best. Both graphs show a large amount of N/A answers due to only 50% of respondents receiving this question due to survey design. Regardless of opinion on which party is best suited for education, @fig-educ_spend_summary makes it clear that Canadians want the government to focus more resources on education.

### Party Variables

Depending on how the respondent answered certain questions, the survey asked which party they voted for in advance, will vote for, are likely to vote for, or would vote for if possible. Only one of these questions was ever asked so we aggregated those columns into a single `voted_for` variable, filtered to the Liberal and Conservative parties. A preview of the `voted_for` variable can be seen in @tbl-personal_var.

```{r, echo=FALSE}
#| label: tbl-party_var
#| tbl-cap: Sample of the Party the Respondent Supports
head(ces21_data, 5) %>% 
  select(voted_for) %>% 
  kable(
    booktabs = TRUE,
    linesep = ""
  ) %>% kable_styling()
```

```{r, echo=FALSE}
#| label: fig-party_summary
#| fig-cap: Distribution of party support

ces21_data %>% ggplot(aes(x=voted_for)) +
  geom_bar() +
  theme_minimal() +
  geom_text(aes(label = after_stat(count)),
            stat = "count",
            position = position_stack(vjust = 0.5),
            colour = "white")
```

@fig-party_summary shows the Liberal party having more support in this dataset. While the Liberal party won the most seats [@citeSeats] it is worth noting that the Conservative party won the popular vote [@citePopularVote].

# Model {#sec-model}

We investigate one model that might explain political support. A logistic regression model using age, education, and income. Logistic regression is well-suited for this problem due to limiting political support to a binary outcome variable. We can estimate the probability of political support based on our predictor variables of age, education, and income.

## Model set-up

Define $y_i$ as the political preference of the respondent and equal to 1 if Liberal and 0 if Conservative. Then $age_i$, $education_i$, $income_i$ are the respective answers of the respondent.

```{=tex}
\begin{align*}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\ 
\text{logit}(\pi_i) &= \beta_0 + \beta_1 \times \text{age}_i + \beta_2 \times \text{education}_i + \beta_3 \times \text{income}_i \\ 
\beta_0 &\sim \text{Normal}(0, 2.5) \\ 
\beta_1 &\sim \text{Normal}(0, 2.5) \\ 
\beta_2 &\sim \text{Normal}(0, 2.5) \\ 
\beta_3 &\sim \text{Normal}(0, 2.5) \\ 
\end{align*}
```
In this logistic regression model, we use Normal priors with mean 0 and standard deviation 2.5 for the regression coefficients $\beta_0$, $\beta_1$, $\beta_2$, and $\beta_3$ which are the default priors for `rstanarm`.

# Results {#sec-results}

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library("rstanarm")

personal_model <-
  readRDS(file = here::here("models/personal_model.rds"))

# topic_model <-
#   readRDS(file = here::here("models/topic_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-personal_modelresults
#| tbl-cap: "Explanatory model of party support based on age, education, and income"
#| warning: false
#| fig-height: 5

library("modelsummary")

tab <- modelsummary::modelsummary(
  list(
    "Support Liberal" = personal_model
  ),
  statistic = "mad",
  fmt = 2,
  output = "kableExtra"
)

tab %>% kable_styling(latex_options = "scale_down")
```

Based on @tbl-personal_modelresults we observe that age has a coefficient of -0.01, indicating age has little effect on party support. Bachelor's degree, Master's degree and Professional degree or doctorate have coefficients of 2.95, 3.26, and 3.06. When compared to lower levels of educations such as completed elementary school, some secondary school, and completed secondary school with coefficients of 2.05, 2.15, and 2.40 it suggests those with higher education are more likely to vote Liberal.

No income, and the income ranges \$30,001 to \$60,000 and \$60,001 to \$90,000 have coefficients -0.61, -0.19, and -0.33. When compared to the higher income ranges \$90,001 to \$110,000, \$110,001, to \$150,000, \$150,001 to \$200,000, and more than \$200,000, which have coefficients -0.54, -0.45, -0.65, and -0.65, it suggests those with higher income are more likely to vote Conservative. Refer to [Appendix -@sec-credibility-intervals] for the credibility intervals of the predictors.

# Discussion {#sec-discussion}

## Conclusion

By analyzing the 2021 CES dataset, we find that age makes no difference for party support, and those with higher education are likely to vote Liberal, while those with higher income are more likely to vote Conservative. We claim that support for the primary parties of Canada is based more on a divide between policy rather than demographics, at least in regards to age, education, and income. This is in contrast to the United States where there are significant differences in political support based on demographics [@citeDemographics].

## Lack of Topical Model {#sec-first-point}

The paper omits a model based on topic variables due to run-time. Due how the data was stored, each variable is a number that was converted to a factor using `as_factor` from `haven` [@citeHaven]. These variables would often have 3 to 5 levels for each factor or more. This lead to large models for the topic model and even the personal model. This was simply not feasible, however the data processing prior to the decision to omit the topic model and simplify the personal model was kept in the paper.

## Model Limitations

Our model has an R2 of 0.031, meaning approximately 3.1% of variance in the outcome variable is explained by our chosen predictor variables. This value is relatively low even in the context of social science, indicating that we might need to look at other factors to fully explain political support. Despite this our model still finds a relationship similar to @fig-educ_party_summary and @fig-econ_party_summary where people concerned with the economy are likely to vote Conservative and those concerned with education are likely to vote Liberal.

Logistic regression assumes a linear relationship between our predictor variables and the log-odds of our outcome variable. This assumption may not hold up, especially when our variables are correlated. For example, there is a correlation between age and education in the sense that there are average ages for finishing levels of education.

## Next Steps

A model using topic variables or a combination of personal variables and topic variables could produce a better R2 than what we obtained. Such a model would be useful for prediction in the upcoming election. Geographical data is also a key part of analyzing voter tendencies, as parties have dominated cities or provinces historically.

Translate the findings into practical recommendations for political campaigns and policymakers. Identify key policy areas and messages that resonate with different demographic groups and regions. Use the models to inform strategic decision-making, resource allocation, and targeted campaigning to maximize electoral success.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional Data Details {#sec-stratification-goals}

## Stratification Goals

### Gender

-   50% men

-   50% women

### Age

-   28% aged 18-34

-   33% aged 35-54

-   39% aged 55 and higher

### Region

-   7% in the Atlantic

-   23% in Quebec

-   38% in Ontario

-   32% in the West

### Language

-   80% French in Quebec, 20% English

-   10% French within the Atlantic

-   10% French nationally

## Response Quality {#sec-response-quality}

As per @citeCES, responses were kept according to the following on page 6 of the codebook:

During the data cleaning process, respondents were categorized based on their most important reason for removal. While respondents might be removed for multiple reasons, the most important reason is most relevant. Reasons for removal are, in order of importance:

1.  Internal survey testing or previews

2.  Ineligible - did not consent to survey

3.  Ineligible - not a Canadian citizen or permanent resident

4.  Ineligible - respondent under 18 years of age

5.  Over quota

6.  Incomplete - did not complete the survey

7.  Duplicate of previous respondent (identified by survey panel ID)

8.  Duplicate of a previous respondent (identified by IP address and the following demographics: year of birth, gender, education level, employment, religion, immigration status)

9.  Speeder (completed the survey in less than 500 seconds, or 8.3 minutes)

10. Postal code-province mismatch

11. Straightliner

12. Failed attention check

13. YOB mismatch

14. Province mismatch

15. Inattentive

16. Initial duplicate (identified by survey panel ID)

The following values are in the final dataset

17. Initial duplicate (identified by IP address and demographics)

18. PES speeders (respondents that took between 6 and 7.12 minutes to complete the PES)

19. Clean complete

# Model Details

## Posterior Predictive Check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Examining how the model fits the data with a posterior prediction check"

pp_check(personal_model) +
  theme_classic() +
  theme(legend.position = "bottom")
```

In @fig-ppcheck we implement a posterior predictive check. The results suggest that the data is consistent with the model predictions. The model adequately captures the observed patterns and variability in the data.

## Predictor Credibility Intervals {#sec-credibility-intervals}

```{r, echo=FALSE, fig.width=10}
#| label: fig-cred_level
#| fig-cap: Credible intervals for predictors of party support
modelplot(personal_model, conf_level = 0.9) +
  labs(x = "90 per cent credibility interval")
```

@fig-cred_level shows small credibility intervals for income categories and age; and larger credibility intervals for education. The coefficients for income remain negative and the coefficients for education remain positive in 90% credibility intervals. This indicates the direction of their remains the same even if the strength changes.

\newpage

# References